---
title: "Analysis of general linear model"
output:
  word_document: default
---
## QUE-
Consider the following model.
yi = θi +θ5 +εi , i = 1, 2, 3
y6 = θ3 +θ7 +ε6
y4 = θ4 +θ6 +ε4
y7 = θ2 +θ8 +ε7
y5 = θ1 +θ7 +ε5
y8 = θ4 +θ8 +ε8
(a) Write the model in Gauss-Markoff setup and find rank of error space and estimation space.
(b) Check the estimability of the following parametric functions.i)θ2 −θ4ii) θ1 +θ2 +θ3 +3θ5 
iii)θ1 −θ2 iv) θ3 +2θ5 −θ1 −2θ8
(c) If Y = (60.2 74.39 77.88 94.75 81.47 99.34 111.86 127.68)
obtain two different solutions to
normal equations and verify that the BLUE of an estimable parametric function is unique
even though two different solutions to normal equations are used.
(d) Obtain an estimate of error variance, BLUE of Xβ and variance-covariance of the BLUE of
Xβ


```{r}
x1=c(1,0,0,0,1,0,0,0)
x2=c(0,1,0,0,1,0,0,0)
x3=c(0,0,1,0,1,0,0,0)
x4=c(0,0,0,1,0,1,0,0)
x5=c(0,0,0,1,0,0,1,0)
x6=c(0,0,1,0,0,0,1,0)
x7=c(0,1,0,0,0,0,0,1)
x8=c(0,0,0,1,0,0,0,1)
X=matrix(c(x1,x2,x3,x4,x5,x6,x7,x8),nrow=8,ncol=8,byrow=T)
X=rbind(x1,x2,x3,x4,x5,x6,x7,x8)
r=qr(X)$rank
r   #Rank of estimation space
n=8
n-r #Rank of error space

```
## To Check Estimability
```{r}
S=t(X)%*%X
r=qr(S)$rank
library(MASS)
G=ginv(S)
H=G%*%S
#To check estimability of theta2-theta4
l1=matrix(c(0,1,0,-1,0,0,0,0),8,1)
t(l1)
zapsmall(t(l1)%*%H)
#To check estimability of theta1+theta2
l2=matrix(c(1,1,0,0,0,0,0,0),8,1)
t(l2)
zapsmall(t(l2)%*%H)
```

## To Find Two Different Solutins To Normal Equations
```{r}
Y=matrix(c(60.2,74.39,77.88,94.75,81.47,99.34,111.86,127.68))
beta_hat=G%*%t(X)%*%Y
z=matrix(c(0,1,0,0.5,0,1,0,4),8,1)
beta_tilda=beta_hat+(diag(8)-H)%*%z

t(l1)%*%beta_hat
t(l1)%*%beta_tilda

```

## To Estimate Error Variance
```{r}
Y_hat=X%*%beta_hat
cbind(Y,Y_hat)
e=Y-Y_hat;e
SSE=t(e)%*%e;SSE
sigma2_hat=SSE/(n-r);sigma2_hat

```

