---
title: "Analysis of one way and two way anova"
output: word_document
date: "2023-01-26"
---

## QUE-
A study was done to compare Pinot Noir wine made in three different regions. Wine samples from
each region were taken and scored by a panel of judges on the wine quality flavor. The higher the
number, the more favorable the rating.
Region 1 3.1 3.5 4.8 3.1 5.5 5.0 4.8 4.3 4.7 4.3 5.1 3.0 4.3 5.5 4.2 3.5 5.7
Region 2 4.3 3.4 5.0 4.1 4.7 5.0 5.0 2.9 5.0
Region 3 3.9 4.5 7.0 6.7 5.8 5.6 4.8 5.5 6.6 5.3 5.7 6.0
(a) Compare the flavor ratings for three different regions graphically using appropriate graph.
(b) Analyze the data using suitable model. Test whether the flavor ratings for different regions
are significantly different.
(c) Obtain residuals and fitted values for this model.

```{r}
n1=17
n2=9
n3=12
n=n1+n2+n3
x1=c(rep(1,n))
x2=c(rep(1,n1),rep(0,n2+n3))
x3=c(rep(0,n1),rep(1,n2),rep(0,n3))
x4=c(rep(0,n1+n2),rep(1,n3))
X=cbind(x1,x2,x3,x4)
S=t(X)%*%X
r=qr(X)$rank
library(Matrix)
r=rankMatrix(S)
r=qr(S)$rank
r=r[1]
library(MASS)
G=ginv(S)
y1=c(3.1,3.5,4.8,3.1,5.5,5.0,4.8,4.3,4.7,4.3,5.1,3.0,4.3,5.5,4.2,3.5,5.7)
y2=c(4.3,3.4,5.0,4.1,4.7,5.0,5.0,2.9,5.0)
y3=c(3.9,4.5,7.0,6.7,5.8,5.6,4.8,5.5,6.6,5.3,5.7,6.0)
Y=matrix(c(y1,y2,y3),n,1)
beta_hat=G%*%t(X)%*%Y

Y_hat=X%*%beta_hat
e=Y-Y_hat
SSE=t(e)%*%e
SSTret=t(beta_hat)%*%t(X)%*%Y-n*mean(Y)^2
TSS=t(Y)%*%Y-n*mean(Y)^2
SST=SSE+SSTret
MSTret=SSTret/(n-r)
MSE=SSE/(n-r)
F=MSTret/MSE
Ftab=qf(0.05,r-1,n-r,lower.tail=FALSE)
Ftab
boxplot(y1,y2,y3)
factor=c(rep("y1",n1),rep("y2",n2),rep("y3",n3))
data=data.frame(Y,factor)
fit=aov(Y~factor)
summary(fit)
TukeyHSD(fit,conf.level=0.95)
R2=(SSTret/TSS)*100  #explained variablity
R2
Unexplained=(SSE/TSS)*100
```



## QUE-
The data in table aside shows the life of electric bulbs from four companies A, B, C and D. Prepare
analysis of variance table to find out whether the lives differ from company to company.
Company Life in Burning hours
A 1020 1010 1030 1000
B 1030 1040 1050 1030 1060
C 990 980 970 960 970 980
D 1040 1050 1030 1070


```{r}
n1=4
n2=5
n3=6
n4=4
n=n1+n2+n3+n4
x1=c(rep(1,n))
x2=c(rep(1,n1),rep(0,n2++n3+n4))
x3=c(rep(0,n1),rep(1,n2),rep(0,n3+n4))
x4=c(rep(0,n1+n2),rep(1,n3),rep(0,n4))
x5=c(rep(0,n1+n2+n3),rep(1,n4))
X=cbind(x1,x2,x3,x4,x5)
S=t(X)%*%X
r=qr(X)$rank
library(Matrix)
r=rankMatrix(S)
r=qr(S)$rank
r=r[1]
library(MASS)
G=ginv(S)
y1=c(1020,1010,1030,1000)
y2=c(1030,1040,1050,1030,1060)
y3=c(990,980,970,960,970,980)
y4=c(1040,1050,1030,1070)
Y=matrix(c(y1,y2,y3,y4),n,1)
beta_hat=G%*%t(X)%*%Y
Y_hat=X%*%beta_hat
e=Y-Y_hat
SSE=t(e)%*%e
SSTret=t(beta_hat)%*%t(X)%*%Y-n*mean(Y)^2
TSS=t(Y)%*%Y-n*mean(Y)^2
SST=SSE+SSTret
MSTret=SSTret/(n-r)
MSE=SSE/(n-r)
F=MSTret/MSE
Ftab=qf(0.05,r-1,n-r,lower.tail=FALSE)
Ftab
boxplot(y1,y2,y3,y4)
factor=c(rep("y1",n1),rep("y2",n2),rep("y3",n3),rep("y4",n4))
data=data.frame(Y,factor)
fit=aov(Y~factor)
summary(fit)
TukeyHSD(fit,conf.level=0.95)
R2=(SSTret/TSS)*100  #explained variablity
R2
Unexplained=(SSE/TSS)*100

```
## QUE-
University of Wisconsin researcher tested the yields of six varieties of Alfalfa on each of four separate fields. The data is given in the following table.
Field Variety
A B C D E F
1 3.22 3.04 3.06 2.64 3.19 2.49
2 3.31 2.99 3.17 2.75 3.40 2.37
3 3.26 3.27 2.93 2.59 3.11 2.38
4 3.25 3.20 3.09 2.62 3.23 2.37
(a) Using appropriate model test whether the yields of six varieties are significantly          different.
(b) Do the yields from different fields differ significantly?
(c) Obtain the BLUEs and estimates of variance of BLUEs of contrast effects in variety and      fields.
(d) Present the variation in average yield due to different varieties and fields                graphically.

```{r}
y1=c(3.22,3.04,3.06,2.63,3.19,2.49)
y2=c(3.31,2.99,3.17,2.75,3.40,2.37)
y3=c(3.26,3.27,2.93,2.59,3.11,2.38)
y4=c(3.25,3.20,3.09,2.62,3.23,2.37)
boxplot(y1,y2,y3,y4)
n=length(y1)+length(y2)+length(y3)+length(y4)
Y=matrix(c(y1,y2,y3,y4),n,1)
field=c(rep("1",6),rep("2",6),rep("3",6),rep("4",6))
m=c("A","B","C","D","E","F")
variety=rep(m,4)
data=data.frame(Y,field,variety)
fit=aov(Y~field+variety,data=data)
fit
summary(fit)


```






