---
title: "Logistic Regression"
author: "Desale Kalyani Bhausaheb"
output: word_document
---
## QUE-
The table below presents the test-firing results for 25 surface-to-air antiaircraft missiles at targets
of varying speed. The result of each test is either a hit (y = 1) or a miss (y = 0).

Test Target Speed x (Knots) y Test Target Speed x (Knots) y
1 400 0 14 330 1
2 220 1 15 280 1
3 490 0 16 210 1
4 210 1 17 300 1
5 500 0 18 470 1
6 270 0 19 230 0
7 200 1 20 430 0
8 470 0 21 460 0
9 480 0 22 220 1
10 310 1 23 250 1
11 240 1 24 200 1
12 490 0 25 390 0
13 420 0
(a) Fit a logistic regression model to the response variable y. Use a simple linear regression model
as the structure for the linear predictor.
(b) Does the model deviance indicate that the logistic regression model from part a is adequate?
(c) Provide an interpretation of the parameter β1 in this model.
(d) Expand the linear predictor to include a quadratic term in target speed. Is there any evidence
that this quadratic term is required in the model?

```{r}
y=c(0,1,0,1,0,0,1,0,0,1,1,0,0,1,1,1,1,1,0,0,0,1,1,1,0)
x=c(400,220,490,210,500,270,200,470,480,310,240,490,420,330,
    280,210,300,470,230,430,460,220,250,200,390)
plot(x,y)
#i)
fit=glm(y~x, family="binomial")
fit
# ii)
summary(fit)   
qchisq(0.05,23,lower.tail = F)
# OR
#library(ResourceSelection)
#hoslem.test(y,fitted(fit),g=10)
# iii)
odds=exp(coef(fit))
odds=exp(cbind(coef(fit),confint(fit)))
odds

fitted=fitted.values(fit)
plot(x,fitted)
```

## QUE-
 A study was performed to investigate new automobile purchases. A sample of 20 families was
selected. Each family was surveyed to determine the age of their oldest vehicle and their total
family income. A follow - up survey was conducted 6 months later to determine if they had actually
purchased a new vehicle during that time period (y = 1 indicates yes and y = 0 indicates no). The
data from this study are shown in the following table.
Income x1 Age x2 y Income x1 Age x2 y
45000 2 0 37000 5 1
40000 4 0 31000 7 1
60000 3 1 40000 4 1
50000 2 1 75000 2 0
55000 2 0 43000 9 1
50000 5 1 49000 2 0
35000 7 1 37500 4 1
65000 2 1 71000 1 0
53000 2 0 34000 5 0
48000 1 0 27000 6 0
(a) Fit a logistic regression model to the data.
(b) Does the model deviance indicate that the logistic regression model from part i) is adequate?
(c) Interpret the model coefficients β1 and β2.
(d) What is the estimated probability that a family with an income of $45,000 and a car that is 5
years old will purchase a new vehicle in the next 6 months?
(e) Expand the linear predictor to include an interaction term. Is there any evidence that this
term is required in the model?
(f ) For the model in part i), find statistics for each individual model parameter.
(g) Find approximate 95% confidence intervals on the model parameters for the logistic regression model from part a).
```{r}
rm(list=ls(all=TRUE))
x11=c(45,40,60,50,55,50,35,65,53,48,37,31,40,75,43,49,37.5,71,34,27)
x1=x11*1000
x2=c(2,4,3,2,2,5,7,2,2,1,5,7,4,2,9,2,4,1,5,6)
y=c(0,0,1,1,0,1,1,1,0,0,1,1,1,0,1,0,1,0,0,0)
fit=glm(y~x1+x2,family="binomial")
fit
summary(fit)
fitted_Model=1/(1+exp(-(-7.047e+00+7.382e-05*x1+9.879e-01*x2)))
odds=exp(coef(fit))
fitted=fitted.values(fit)
fitted_Model=1/(1+exp(-(-7.047e+00+7.382e-05*45000+9.879e-01*5)))
Model_2=glm(y~x1+x2+x1:x2,family="binomial")
summary(Model_2)
Model_3=glm(y~x1*x2,family="binomial")

```











