---
title: "Polynomial Regression"
output: word_document
---

## QUE-
 Table 3          Tensile
Hardwood          Strength
Conc. (xi) in      (yi) in psi
%

1 6.3
1.5 11.1
2 20
3 24
4 26.1
4.5 30
5 33.8
5.5 34
6 38.1
6.5 39.9
7 42
8 46.1
9 53.1
10 52
11 52.5
12 48
13 42.8
14 27.8
15 21.9
given Table  presents the data concerning the strength of Kraft paper and the % of hardwood in the
batch of pulp from which the paper was produced. Show that the quadratic model appears to
provide a good fit to the relationship between tensile strength and hardwood concentration.


```{r}
x=c(1,1.5,2,3,4,4.5,5,5.5,6,6.5,7,8,9,10,11,12,13,14,15)
y=c(6.3,11.1,20,24,26.1,30,33.8,34,38.1,39.9,42,46.1,53.1,
    52,52.5,48,42.8,27.8,21.9)
plot(x,y)
Model_1=lm(y~x+I(x^2))
summary(Model_1)
# centering the data
x_1=x-mean(x)
x_2=(x-mean(x))^2
Model_2=lm(y~x_1+I(x_2))
summary(Model_2)
#library(car)
#vif(Model_1)
#vif(Model_2)
```

## QUE-
Fit the regression model as specified below for the data given in Table 2
y = β0 +β1x1 +β2x2 +β11x2^2+β22x2^2+β12x1x2
(a) Compute t-statistic for each model parameter and conclude on the significance of corresponding regression parameter.
(b) Use extra sum of squares method to evaluate the contribution of all the quadratic terms x1^2,x2^2
and x1x2 in regression sum of squares for the regression model fitted in (a). Also test its significance.
(c) Plot the residuals versus fitted values. Also have the Box plot of residuals to check the existence of any of the outlying observations.
26 1 1 
24 1 1
175 1.5 4 
160 1.5 4 
163 1.5 4 
55 0.5 2 
62 1.5 5 
100 0.5 3 
26 1 1.5 
30 0.5 1.5 
70 1 2.5 
71 0.5 2.5


```{r}
y=c(26,24,175,160,163,55,62,100,26,30,70,71)
x1=c(1,1,1.5,1.5,1.5,0.5,1.5,0.5,1,0.5,1,0.5)
x2=c(1,1,4,4,4,2,5,3,1.5,1.5,2.5,2.5)
data=data.frame(y,x1,x2)
data
plot(data)
fit=lm(y~x1+x2+I(x1^2)+I(x2^2)+I(x1*x2))
fit  
#   a)
summary(fit)
#  b) 
# FM
fit1=lm(y~x1+x2+I(x1^2)+I(x2^2)+I(x1*x2))
fit2=lm(y~x1+x2)  
anova(fit1,fit2)
qf(0.05,3,6,lower.tail = F)
# manually
anova(fit1)
anova(fit2)
SSRes_FM=655.8  
SSRes_RM=14472.5 
SSRes_RM
SSHo=SSRes_RM-SSRes_FM
SSHo
MSRes=10.45499^2
F_o=(SSHo/3)/MSRes
F_o

# c)
# To check constant variance assumption
par(mfrow=c(1,2))
plot(fit1$fitted.values,fit1$residuals)

#library(lmtest)
#bptest(fit1)
# To detect outliers
boxplot(fit1$residuals,outline=T)
r=rstudent(fit1)
r
r[abs(r)>2]
boxplot(r,outline=T)
# Test for normality
#shapiro.test(r)
#library(nortest)
#ad.test(r)
```



